---
title: "project"
author: "project"
date: "2024-11-13"
output: html_document
---
```{r}
#Installing Required packages
install.packages('DT')
install.packages('psych')
install.packages('tidyr')
install.packages("shiny")

```

```{r}
#Libraries
library(psych)
library(tidyr)
library(shiny)
library(dplyr)
library(data.table)
library(DT)
library(ggplot2)
```
*Data Loading and Preprocessing*
```{r}
# Load the dataset
data_cleaned_encoded <- fread("C:\\Users\\sivac\\Desktop\\merged_data_final.csv")
```
```{r}
#Displaying some data
head(data_cleaned_encoded)
```

```{r}
dim(data_cleaned_encoded) #Dimensions
```

```{r}
describe(data_cleaned_encoded) #Descriptive statistics
```

```{r}
str(data_cleaned_encoded)
```

```{r}
#checking for missing values,if any
na_value <- sapply(data_cleaned_encoded, function(x) sum(is.na(x)))
b_value <- sapply(data_cleaned_encoded, function(x) sum(x == ""))
df_value <- data.frame(
  Column = names(na_value),
  NA_Count = na_value,
  Blank_Count = b_value
)

df_value
```

```{r}
sampled_data <- data_cleaned_encoded[sample(1:nrow(data_cleaned_encoded), 10), ] #Displaying some random samplesfrom the data
sampled_data
```
*Exploratory Data Analysis (EDA)*
```{r}
genres_split <- strsplit(as.character(data_cleaned_encoded$genres), "\\|")

all_genres <- unlist(genres_split)

unique_genres <- unique(all_genres)

print(unique_genres)
```
*Feature Engineering*

*Rating Normalization*
```{r}
# Normalize genres and ratings for better recommendations
data_cleaned_encoded <- data_cleaned_encoded %>%
  group_by(userId, title, genres) %>%
  summarize(rating = mean(rating, na.rm = TRUE), .groups = "drop") %>%
  group_by(userId) %>%
  mutate(rating_normalized = rating - mean(rating, na.rm = TRUE)) %>%
  ungroup()

```
*To account for user bias, rating normalization was used by adjusting each movie rating based on the user’s average. This helps balance out differences in how users rate, whether they tend to score things higher or lower than others. By normalizing the ratings, we ensure a fairer comparison of preferences, which is essential for making personalized recommendations.*

*Data Visualization*
```{r}
# Process the dataset to extract top genres
top_genres <- data_cleaned_encoded %>%
  separate_rows(genres, sep = "\\|") %>%
  count(genres, sort = TRUE) %>%
  top_n(10, n)

# Plot the top genres as a bar chart
ggplot(top_genres, aes(x = reorder(genres, n), y = n, fill = genres)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme_minimal() +
  labs(
    title = "Top 10 Movie Genres by Count",
    x = "Genres",
    y = "Count",
    subtitle = "Top genres from the dataset"
  ) +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 20, margin = margin(b = 25))
  )+
  geom_text(aes(label = n), hjust = -0.1, size = 4)
```
*Popular genres like Action and Comedy highlight key clusters of movies, which item-based filtering can use to find similarities. This ensures recommendations align with user preferences for frequently co-rated genres.*
```{r}

# Explore Top Genres with percentages
top_genres <- data_cleaned_encoded %>%
  separate_rows(genres, sep = "\\|") %>%
  count(genres, sort = TRUE) %>%
  top_n(10, n) %>%
  mutate(percentage = round(n / sum(n) * 100, 1))

# Plot the Top Genres as a Pie Chart with Labels
ggplot(top_genres, aes(x = "", y = n, fill = genres)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  geom_text(aes(label = paste0(percentage, "%")),
            position = position_stack(vjust = 0.5), size = 4) +
  theme_minimal() +
  labs(title = "Top Movie Genres by percentage", y = "Count", x = "") +
  theme(axis.text.x = element_blank(),
        plot.title = element_text(size = 16, margin = margin(b = 20)),
        legend.position = "right")

```
*The proportional view of genres helps focus item-based filtering on dominant categories. It ensures the algorithm prioritizes well-represented genres, improving the relevance of suggestions.*
```{r}
Top_25 <-  data_cleaned_encoded%>%
  group_by(title) %>%
  summarise(Rating_Count = n()) %>%
  arrange(desc(Rating_Count)) %>%
  slice_head(n = 25)

# View the result
print(Top_25)

```

```{r}
ggplot(Top_25, aes(x = reorder(title, Rating_Count), y = Rating_Count, fill = Rating_Count)) +
  geom_bar(stat = "identity") +
  coord_flip() +  # Creates a horizontal bar chart
  labs(title = "Top 25 Rated Movies",
       x = "Top 25 Movie Names",
       y = "Total Ratings") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, margin = margin(b = 20)),
        axis.text.y = element_text(size = 10),
        axis.text.x = element_text(size = 10)) +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_gradient(low = "lightgrey", high = "black")
```
*Frequently rated movies like Pulp Fiction serve as anchors for similarity calculations in item-based filtering. These titles help the model identify other movies co-rated by users with similar preferences.*
```{r}
distinct_movie_count <-  data_cleaned_encoded %>%
  summarise(total_distinct_movies = n_distinct(title))

# Print the number of distinct movies
print(distinct_movie_count)
```

```{r}
#Filter the data for the movie 'Pulp Fiction'
pulp_fiction_ratings <- data_cleaned_encoded %>%
  filter(title == "Pulp Fiction (1994)") %>%
  count(rating)

#Plot the user ratings using a horizontal bar chart
ggplot(pulp_fiction_ratings, aes(x = factor(rating), y = n, fill = factor(rating))) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "User Ratings of the Movie 'Pulp Fiction'",
       x = "Rating",
       y = "Number of Users") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, margin = margin(b = 20), hjust = 0.5),
        axis.text.y = element_text(size = 10),
        axis.text.x = element_text(size = 10)) +
  scale_fill_brewer(palette = "Blues")

```

```{r}

#Count the number of ratings given by each userId
highest_rating_user <- data_cleaned_encoded %>%
  group_by(userId) %>%
  summarise(Rating_Count = n()) %>%
  arrange(desc(Rating_Count)) %>%
  slice_head(n = 10)

# Print the userId and Rating_Count of the user with the most ratings
print(highest_rating_user)

ggplot(highest_rating_user, aes(x = reorder(as.factor(userId), Rating_Count), y = Rating_Count, fill = Rating_Count)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top 10 Users by Number of Ratings Given",
       x = "User ID",
       y = "Number of Ratings") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, margin = margin(b = 20), hjust = 0.5),
        axis.text.y = element_text(size = 10),
        axis.text.x = element_text(size = 10)) +
  scale_fill_gradient(low = "lightblue", high = "darkblue")

```

```{r}
# Average Rating by Genre
avg_genre_rating <- data_cleaned_encoded %>%
  separate_rows(genres, sep = "\\|") %>%
  group_by(genres) %>%
  summarise(average_rating = mean(rating, na.rm = TRUE)) %>%
  arrange(desc(average_rating))

# Bar chart
ggplot(avg_genre_rating, aes(x = reorder(genres, average_rating), y = average_rating, fill = genres)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme_minimal() +
  labs(title = "Average Rating by Genre",
       x = "Genres",
       y = "Average Rating") +
  theme(legend.position = "none")
```
*This bar chart shows the average ratings for different genres, highlighting that genres like Film-Noir and Documentary receive higher average ratings. These genres can be prioritized in recommendations to align with user preferences for high-quality content.*
```{r}
# Top Movies by Average Rating (Minimum 50 Ratings)
top_movies <- data_cleaned_encoded %>%
  group_by(title) %>%
  summarise(average_rating = mean(rating, na.rm = TRUE), count = n()) %>%
  filter(count >= 50) %>%
  arrange(desc(average_rating)) %>%
  slice_head(n = 10)

# Bar chart
ggplot(top_movies, aes(x = reorder(title, average_rating), y = average_rating, fill = average_rating)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme_minimal() +
  labs(title = "Top Movies by Average Rating (Min. 50 Ratings)",
       x = "Movie Titles",
       y = "Average Rating") +
  scale_fill_gradient(low = "lightblue", high = "darkblue") 

```
*This chart showcases the top movies by average rating (with a minimum of 50 ratings), featuring highly acclaimed titles like Fight Club and Pulp Fiction. These popular and well-rated movies act as benchmarks for similarity-based recommendations, guiding users toward high-quality content.*
```{r}
# Density Plot for Normalized Ratings
ggplot(data_cleaned_encoded, aes(x = rating_normalized)) +
  geom_density(fill = "blue", alpha = 0.4) +
  theme_minimal() +
  labs(title = "Distribution of Normalized Ratings",
       x = "Normalized Ratings",
       y = "Density")
```
*This density plot shows the distribution of normalized ratings, centered around zero, indicating effective bias correction for individual user tendencies. This balanced normalization ensures fairer comparisons across users, which is critical for generating accurate recommendations*

*Model Training and Evaluation*
```{r}
calculate_similarity <- function(selected_movies, data,  top_n = 3) {
  if (length(selected_movies) == 0 || nrow(data) == 0) {
    return(data.frame(title = "No movies selected or insufficient data.", similarity = NA, genres = NA))
  }

 # Filter dataset to include only selected movies
  filtered_data <- data %>% filter(title %in% selected_movies)

 # Check if the filtered dataset contains any movies
  if (nrow(filtered_data) == 0) {
    return(data.frame(title = "Selected movies not found in the dataset.", similarity = NA, genres = NA))
  }

  # Extract unique genres from the selected movies
  selected_genres <- unique(unlist(strsplit(filtered_data$genres, "\\|")))

  # Calculate similarity based on genre overlap and normalize by union of genres
  data <- data %>%
    rowwise() %>%
    mutate(similarity = length(intersect(strsplit(genres, "\\|")[[1]], selected_genres)) /
                       length(union(strsplit(genres, "\\|")[[1]], selected_genres))) %>%
    ungroup()

  # Generate recommendations by filtering movies, sorting by similarity and normalized rating
  recommendations <- data %>%
    filter(!title %in% selected_movies & similarity > 0) %>%
    arrange(desc(similarity), desc(rating_normalized)) %>%
    distinct(title, .keep_all = TRUE) %>%
    select(title, similarity, genres) %>%
    head(top_n)

  # If no recommendations found
  if (nrow(recommendations) == 0) {
    return(data.frame(title = "No recommendations found.", similarity = NA, genres = NA))
  }

  return(recommendations)
}

```

```{r}
calculate_lift_score <- function(recommendations, data) {
  if (nrow(recommendations) == 0) {
    return(0)
  }

  relevant_movies <- unique(recommendations$title)
  all_movies <- unique(data$title)

  # Precision of recommendations
  precision <- length(relevant_movies) / length(all_movies)

  # Random precision (baseline)
  random_precision <- 1 / length(all_movies)

  # Lift Score
  lift_score <- ifelse(random_precision > 0, precision / random_precision, 0)
  return(lift_score)
}

```

```{r}
# Load required library
library(data.table)

# Read x_data with proper options
x_data <- fread("C:\\Users\\sivac\\Desktop\\x_data (1).csv", quote = "", fill = TRUE)

# Read y_data with proper options
y_data <- fread("C:\\Users\\sivac\\Desktop\\y_data (1).csv", quote = "", fill = TRUE)

parse_movie_list <- function(movie_string) {
  # Step 1: Remove leading/trailing brackets and quotes
  movie_string <- gsub("^\\[|\\]$", "", movie_string)  # Remove square brackets at start/end
  movie_string <- gsub("'", "", movie_string)         # Remove single quotes
  movie_string <- gsub('"', '', movie_string)         # Remove double quotes
  
  # Step 2: Handle leading or trailing unmatched brackets
  movie_string <- gsub("^\\[", "", movie_string)  # Remove unmatched leading brackets
  movie_string <- gsub("\\]$", "", movie_string)  # Remove unmatched trailing brackets
  
  # Step 3: Split into individual titles
  movie_list <- strsplit(movie_string, ",\\s*")[[1]]  # Split on commas with optional spaces
  
  # Step 4: Trim spaces and artifacts around each movie title
  movie_list <- trimws(movie_list)  # Remove leading/trailing spaces
  return(movie_list)
}

# Apply parsing to x_data and y_data
x_data_parsed <- lapply(x_data[[1]], parse_movie_list) 
y_data_parsed <- lapply(y_data[[1]], parse_movie_list)  

# Check the parsed results
print(x_data_parsed[[1]])
print(y_data_parsed[[1]])

```

```{r}
# Initialize an empty list to store scores and recommendations
results <- list()

# Loop through all sets
for (i in seq_along(x_data_parsed)) {
  # Generate recommendations
  recommendations <- calculate_similarity(x_data_parsed[[i]], data_cleaned_encoded, top_n = 3)
  
  # Calculate score
  overlap_actual <- length(intersect(recommendations$title, y_data_parsed[[i]]))
  overlap_input <- length(intersect(recommendations$title, x_data_parsed[[i]]))
  score <- (overlap_actual - overlap_input) / 3

  # Store results
  results[[i]] <- list(
    Input = x_data_parsed[[i]],
    Actual = y_data_parsed[[i]],
    Recommendations = recommendations$title,
    Score = score
  )
}
# Calculate average score
scores <- sapply(results, function(res) res$Score)
average_score <- mean(scores, na.rm = TRUE)

# Print the average score
print(paste("Average Score:", average_score))

# Print results for all rows
#for (i in seq_along(results)) {
 # print(paste("Set", i, "Results:"))
 # print(results[[i]])
#}
```

```{r}
# Process the first two lists of movies from x_data_parsed
first_two_movies <- x_data_parsed[1:2]

# Function to generate and print recommendations for each list
generate_recommendations <- function(movie_list, index) {
  recommendations <- calculate_similarity(movie_list, data_cleaned_encoded, top_n = 3)
  
  # Print recommendations
  cat("\nRecommendations for List", index, ":\n")
  if (nrow(recommendations) > 0) {
    print(recommendations)
  } else {
    cat("No recommendations found.\n")
  }
}

# Iterate over the first two lists and generate recommendations
for (i in seq_along(first_two_movies)) {
  generate_recommendations(first_two_movies[[i]], i)
}
```

```{r}
library(shiny)

# Define UI
ui <- fluidPage(
  titlePanel("Movie Recommendation System"),  # Title of the app
  
  sidebarLayout(
    sidebarPanel(
      textInput("movies", "Enter Movies (comma-separated):", 
                placeholder = "e.g., Inception, Titanic, Avatar"),  # Input for movies
      actionButton("recommend", "Get Recommendations"),  # Button to fetch recommendations
      helpText("Enter one or more movie names separated by commas.")
    ),
    
    mainPanel(
      tabsetPanel(
        tabPanel("Recommendations", 
                 h3("Recommended Movies"),
                 tableOutput("recommendations")),  # Output table for recommendations
        
        tabPanel("Input",
                 h3("Input Movies"),
                 tableOutput("input_movies")),  # Output table for input movies
      )
    )
  )
)

```

```{r}
server <- function(input, output) {
  observeEvent(input$recommend, {
    # Parse the input movies
    input_movies <- unlist(strsplit(input$movies, ",\\s*"))  # Split by commas and trim spaces
    
    # Validate input
    if (length(input_movies) > 0) {
      # Generate recommendations
      recommendations <- calculate_similarity(input_movies, data_cleaned_encoded, top_n = 3)
      
      # Render recommendations
      output$recommendations <- renderTable({
        if (nrow(recommendations) > 0) {
          data.frame(Recommended_Movies = recommendations$title)
        } else {
          data.frame(Error = "No recommendations available for the entered movies.")
        }
      })
      
      # Render input movies
      output$input_movies <- renderTable({
        data.frame(Input_Movies = input_movies)
      })
      
      # Optional: Render evaluation metrics or plots
      output$metrics <- renderText({
        paste("Number of Recommendations:", nrow(recommendations))
      })
    } else {
      output$recommendations <- renderTable({
        data.frame(Error = "Please enter one or more movie names.")
      })
    }
  })
}

```

```{r}
shinyApp(ui = ui, server = server)
```



